---
title: 'Data Analysis 3: Week 9'
author: "Alexey Bessudnov"
date: "11 March 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(cache = TRUE)

```
Plan for today:

1. Assignment 4: solutions.
2. Assignment 5.
3. Exercises on conditional statements and iteration.


Exercises.

1. Check if a number is even. If it is print "Even". If it isn't print "Odd". If it's 0 print "Zero".

    ```{r}
x <- 3

if (x == 0) {
  print("Zero")
} else {
    if (x %% 2 == 0) {
      print("Even")
    } else {
        print("Odd")
    }
  }

    ```

2. Write a for loop finding the largest element of a numeric vector and print it on the screen.

    ```{r}
x <- c(3, 5, 8, -2)
max(x)

for (i in 1:3) {
  print(i)
}

for(i in 1:length(x)) {
  print(x[i])
}

max_x <- x[1]
for (i in 2:length(x)) {
  if (x[i] > max_x) {max_x <- x[i]}
}
max_x

    ```

3. Write a for loop finding the largest element of a numeric matrix and print it on the screen.

    ```{r}
x <- matrix(c(1, -3, 7, 4, 10, 7, -5, 0), nrow = 4)
x
max_x <- x[1, 1]
for (i in 1:dim(x)[1]) {
  for (j in 1:dim(x)[2]) {
      if (x[i,j] > max_x) {max_x <- x[i,j]}
  }
}
max_x

    ```

4. Write a while loop finding the largest element of a numeric vector and print it on the screen.

    ```{r}
i <- 1 
while (i < 10) {
  print(i)
  i <- i + 1
}

x <- c(3, 5, 8, -2)

max_x <- x[1]
i <- 2
while (i <= length(x)) {
    if (x[i] > max_x) {max_x <- x[i]}
  i <- i + 1
}
max_x

    ```

5. x is a vector with whole numbers (zero and positive integers).  If the largest even element of x is smaller than the largest odd element of x, all even elements of x are replaced by 0s. Otherwise all odd elements of x are replaced by 0s. For example, if x = {7; 1; 3; 2; 14; 5; 9; 6} the output should be [0; 0; 0; 2; 14; 0; 0; 6].

    ```{r}
x <- c(7, 1, 3, 2, 14, 5, 9, 6)
max <- 0
for (i in seq_along(x)) {
                if (x[i] > max) {max <- x[i]}
}
for (i in seq_along(x)) {
        if (max %% 2 == 0) {
                if (x[i] %% 2 != 0) {x[i] <- 0}
        }
        else {
                if (x[i] %% 2 == 0) {x[i] <- 0}
        }
}
x
    ```

6. x is a vector with whole numbers. Write a programme that counts the number of pairs of the elements of x where the sum can be divided by 12 without a remainder.

For example, for x = {8; 10; 14; 7; 13; 5; 30; 9; 6} then the answer is 3 ((10, 14); (7, 5); (30, 6)).

    ```{r}
x <- c(8, 10, 14, 7, 13, 5, 30, 9, 6)
k <- 0
for (i in 1:(length(x)-1)) {
        for (j in (i+1):length(x)) {
                if ((x[i] + x[j]) %% 12 == 0) {
                         k <- k + 1
                         print(c(x[i], x[j]))
                }
        }
}
k
    ```


7. Write a function to calculate the factorial (i.e. 5! = 1x2x3x4x5). Note that 0! = 1, and for the negative numbers the factorial is not defined. (Hint: use recursion.)

```{r}
myFactorial <- function(x){
  if (x == 0) {return(1)}
  x * myFactorial(x-1)
}
myFactorial(4)
```

8. Use this function to calculate the factorial for the numbers from 1 to 10. (Hint: use map()).

```{r}
# this is not going to work because the function has not been vectorised
myFactorial(1:10)
# a for() loop
output <- numeric(10)
for (i in 1:10) {
  output[i] <- myFactorial(i)
}
output
# with map()
library(tidyverse)
mapply(myFactorial, 1:10)
map_dbl(1:10, myFactorial)
```


9. Open the indresp file for wave 8 and keep two variables: sex (h_sex_dv) and total net personal income (h_fimnnet_dv). You may need to clean sex and retain only values for male and female. Keep only cases with non-missing values for sex and income.

a) regress income on sex (i.e. find the difference between mean income for men and women). Store the estimate for the regression coefficient.
b) Now regress income on sex 100 times, each time with a random sample of 1,000 people. Store 100 regression coefficients in a vector and illustrate their distribution graphically.
c) Now do the same, changing the sample size to 100 people, and compare two distributions.

```{r}
df <- read_tsv("data/UKDA-6614-tab/tab/ukhls_w8/h_indresp.tab") %>%
        select(h_sex_dv, h_fimnnet_dv)
df <- df %>%
        mutate(h_sex_dv = ifelse(h_sex_dv == 1, "male",
                      ifelse(h_sex_dv == 2, "female", NA))) %>%
        filter(!is.na(h_sex_dv)) %>%
        filter(!is.na(h_fimnnet_dv))
est <- lm(h_fimnnet_dv ~ h_sex_dv, df)$coefficients[[2]]
output <- numeric(100)
for (i in 1:100) {
  df1 <- df %>% sample_n(1000)
  output[i] <- lm(h_fimnnet_dv ~ h_sex_dv, df1)$coefficients[[2]]
}
output
sexCoef <- map(1:100, ~ df %>% sample_n(100)) %>%
        map_dbl(~ lm(h_fimnnet_dv ~ h_sex_dv, data = .)$coefficients[2])
enframe(sexCoef) %>%
        ggplot(aes(x = value)) +
        geom_histogram() +
        geom_vline(xintercept = est, colour = "red")
```





