---
title: "Data Analysis in Social Science  (18 January)"
output:
  html_document:
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(cache = TRUE)



```

## Class on 18 January

Today we will work with the data set USArrests that is a build in data set in R.

```{r}
data(USArrests)
head(USArrests)
```

## Your tasks for today.

1. Extract a vector with the names of the states.

```{r}
USArrests
str(USArrests)
stateNames <- row.names(USArrests)
stateNames
class(stateNames)

USArrests$stateNames <- row.names(USArrests)
USArrests
```

2. Create a new data set that only includes states with the murder rate > 10.

```{r}
USArrests
USArrests[, 5]
USArrests[USArrests$Murder > 10,]
subset(USArrests, USArrests$Murder > 10)

```

3. Create a new data set that only includes the first 15 states in the original data set.

```{r}
USArrests[1:15,]
```

4. Knit your document as pdf, html and Word files.

5. Change the document settings so that the code is not displayed, and messages and warnings are not shown. Set cache to TRUE. When do you want to use cache? Can you change these settings for individual chunks of code?

