---
title: "Data Analysis in Social Science  (8 February)"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(vroom)
```

Exercises.

1) Calculate the average age of parents (separately for women and men) of children born between waves 9 and 10.

    ```{r}
# Read the data for newborn children (born between waves 9 and 10).
# Observations are at the parent level.
    
newborn10 <- vroom("UKDA-6614-tab/tab/ukhls_w10/j_newborn.tab")

# Read the data with people's age.
    
indall10 <- vroom("UKDA-6614-tab/tab/ukhls_w10/j_indall.tab")

# Use semi_join (filtering join) to keep only new parents in the data set.

indall10 %>%
  semi_join(newborn10, by = "pidp") %>%
  # split by gender
  group_by(j_sex) %>%
  # filter out missing ages
  filter(j_age_dv > 0) %>%
  # calculate mean age
  summarise(
    meanAge = mean(j_age_dv, na.rm = TRUE)
  )
    ```

2) Split the table by ethnic group for the following groups: White British, Indian, Pakistani, Bangladeshi, and African.

    ```{r}
# ethnicity is recoded in the cross-wave file
stable <- vroom("UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab",
                col_select = c(pidp, racel_dv))

indall10 %>%
  semi_join(newborn10, by = "pidp") %>%
  # join with ethnicity data
  left_join(stable, by = "pidp") %>%
  # keep only some ethnic groups
  filter(racel_dv %in% c(1, 9, 10, 11, 15)) %>%
  # group by ethnicity and sex
  group_by(racel_dv, j_sex) %>%
  # filter out missing ages
  filter(j_age_dv > 0) %>%
  # calculate mean age
  summarise(
    meanAge = mean(j_age_dv, na.rm = TRUE)
  )

# alternatively, you can us ethe variable j_ethn_dv in indall10

indall10 %>%
  semi_join(newborn10, by = "pidp") %>%
  filter(j_ethn_dv %in% c(1, 9, 10, 11, 15)) %>%
  # group by ethnicity and sex
  group_by(j_ethn_dv, j_sex) %>%
  # filter out missing ages
  filter(j_age_dv > 0) %>%
  # calculate mean age
  summarise(
    meanAge = mean(j_age_dv, na.rm = TRUE)
  )
    ```

3) Produce a data table that includes only twins born between waves 9 and 10. How does the age of their parents compare with the age of parents of non-twin children?

    ```{r}
twins <- newborn10 %>%
      # find parents with more than one new child
      group_by(pidp) %>%
      # create a variable with the number of children per parent
      mutate(n_children = n()) %>%
      # keep only those with > 1 children
      filter(n_children > 1) %>%
      select(pidp, n_children)

# repeat as in ex.1 but only for the parents of twins.
    
indall10 %>%
  semi_join(twins, by = "pidp") %>%
  # split by gender
  group_by(j_sex) %>%
  # filter out missing ages
  filter(j_age_dv > 0) %>%
  # calculate mean age
  summarise(
    meanAge = mean(j_age_dv, na.rm = TRUE)
  )

    ```

