---
title: "Data Analysis in Social Science  (1 February)"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tidyverse)
dfStable <- read_tsv("UKDA-6614-tab/tab/ukhls_wx/xwavedat.tab")
```

This is a cross-wave data file with stable characteristics of individuals. See the codebook at https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/search/datafile/xwavedat.

Exercises.

1. Select the variables for: sex (derived), date of birth (doby_dv), ethnic group (racel_dv). Also keep the cross-wave identifier (pidp) and the sample origin variable (memorig).

    ```{r}

    ```

2. Filter the data to keep:
        a) men only.
        b) people born before 1950 and after 1975.
        c) men of Pakistani origin born in 1958 or 1982. 


    ```{r}

    ```

3. Recode birth year into cohorts (a new variable): the G.I. Generation (born before 1924), Silent Generation (1925-42), Baby Boomers (1943-65), Generation X (1966-1980), Millenials (1981-99), Generation Z (2000-). (The years are approximate.)

    ```{r}

    ```

4. Recode ethnicity into the following groups: white British, Other White, Indian, Pakistani, other. (This classification doesn't make much sense, but we're doing this as an exercise).


    ```{r}

        
    ```


5. Count the number of people belonging to different ethnic groups (and produce percentages). 

    ```{r}

    ```

6. Summarise the proportion of white British by generation.

    ```{r}

    ```
