---
title: "Week 2"
author: "Jack Hooper"
date: "25/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library('vroom')
```


## The Understanding Society questions

1. What is the difference between **indresp** and **indall** files in UndSoc? What kind of information can you find in each?

 * **indresp** files contain only information from adult interviews - individual respondants 
* **indall** files contain all information from adults, children and non-respondants
* Vroom can make mistakes but is fast
* Most of the time use **indresp**
```{r}
a_indresp <- vroom("UKDA-6614-tab/tab/ukhls/a_indresp.tab")
a_indall <- vroom("UKDA-6614-tab/tab/ukhls/a_indall.tab")
head(a_indresp)
head(a_indall)
```


1.b. What kind of information can you find in **xwavedat** files?

* Time constant data
* dv means variable is derived 
* Month and data of DOB not in this release as it is sensitive
* Supposed to be anonymous
* m is maternal information
* wave specific files contain this info but more reliable to use this file and merge with wave specific file

```{r}
xwavedat <- vroom("UKDA-6614-tab/tab/ukhls/xwavedat.tab")

head(xwavedat)
colnames(xwavedat)
```


2. What is the structure of **egoalt** files?

* Kin and other relationships between pairs of individuals in the household
* Links people together in the household (their relationships)
* *pidp* is the individual indentifier across waves
* in this file the individual may have a number of relationships
* each row = 1 relationship
* each relationship appears twice (X with Y and Y with X)
* *hidp* is household identifier
* household identifier is wave specific as people move houses and split and move in and die
* BHPS became part of understanding society dataset
* for some people this data spans 30 years (smaller sample size)


https://www.understandingsociety.ac.uk/documentation/mainstage/dataset-documentation/search/datafile/egoalt
```{r}
a_egoalt <- read.table("UKDA-6614-tab/tab/ukhls/a_egoalt.tab",header=T)
head(a_egoalt)
str
```


* 18 variables *

3. In which files can you find the data on education / highest qualification?

* *qfedhi* contains the highest level of qualification and can be found in *indresp* file
* more variables can be found in the *indresp* file
* also qfhigh_dv - highest ever reported education level
* hiqual_dv

* important to read notes to understand
* Universe shows you conditions req for people to be included

* In longitudinal studies you don't want to repeat the questions but you want to record if there have been changes. You do need to collect data for new people.

4. What kind of information can you find in the **hhresp** and **hhsamp** files?

* **hhsamp** files contain information about the household (household level qs) and is a non harmonised file - *technical data* languages spoken, duration of interview, number of people over 16
* **hhresp** files contain information about the household (household level qs)

5. What kind of information can you find in the **newborn** files?
* DOB, birth weight, sex

6. What is the pidp variable? Why is it important?

* Individual identifier (constant over time)

7. Tell us about the individual with pidp 280165 (sex, age, where they live, etc.).

```{r}
k_indresp <- vroom("UKDA-6614-tab/tab/ukhls/k_indresp.tab")
k_indresp$k_sex_dv[k_indresp$pidp==280165]
k_indresp$k_ppsex[k_indresp$pidp==280165]
xwavedat$bornuk_dv[xwavedat$pidp==280165]
```
* The respondant is female (2 in sex_dv variable)
* The respondant has a male partner when interviewed (1 ppsex)
* She was born in the UK




8. Explore the structure of the questionnaires for wave 11 of UndSoc. 

```{r}
