---
title: "Week 6. Tasks/exercises for 22 February 2022 (visualising data)"
output:
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vroom)
```

## Data visualisation
Communicate statistical information better through graph over numbers.

1. Read the indresp file from Wave 11 and keep the following variables: pidp, derived sex and age , and net personal income (fimnnet_dv). Visualise the distribution of income with a histogram, a density plot and a box plot.

```{r q1}
k_indresp <- vroom("UKDA-6614-tab/tab/ukhls/k_indresp.tab",
            col_select = c(pidp, k_sex_dv, k_age_dv, k_fimnnet_dv))

#always explore variable first

k_indresp %>%
  pull(k_fimnnet_dv) %>% summary()

k_indresp %>%
  ggplot(mapping =aes(x=k_fimnnet_dv))+
  geom_histogram()

# set limits and change bins
k_indresp %>%
  ggplot(mapping =aes(x=k_fimnnet_dv))+
  geom_histogram(bins = 300)+
  xlim(-1000, 10000)

# take log due to extremes

k_indresp %>%
  ggplot()+
  geom_density(mapping =aes(x=k_fimnnet_dv))

k_indresp %>%
  ggplot()+
  geom_boxplot(mapping =aes(x=k_fimnnet_dv))

```

2. Visualise the distribution of sex with a bar chart. Create a variable for age groups and visualise the distribution by age, both as a continuous and categorical variable.

```{r q2}
k_indresp %>%
  count(k_sex_dv)
k_indresp <- k_indresp %>%
  mutate(sex = case_when(
    k_sex_dv == 1 ~ 1,
    k_sex_dv == 2 ~ 2
  ))

k_indresp %>%
  count(sex)

k_indresp %>%
  ggplot(mapping=aes(sex))+
  geom_bar()

```

3. Think of the best way to visualise the relationship between sex and income.

```{r q3}
```

4. Think of the best way to visualise the relationship between age and income.

```{r q4}
```

5. Visualise the joint distribution of age, sex and income.

```{r q5}
```