---
title: "Week 5. Tasks/exercises for 15 February 2022 (reshaping data)"
output:
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vroom)
```

Any data table can be changed to three columns:
ID, column, value
(pivot_longer)

## Reshaping data

1. Join the data from waves 9 to 11 from the **indresp** files and reshape to the long format. Include the variables for age, sex and life satisfaction (sclfsato).


```{r q1}
w9 <- vroom("UKDA-6614-tab/tab/ukhls/i_indresp.tab",
            col_select = c(pidp, i_sex_dv, i_age_dv, i_sclfsato))
w10 <- vroom("UKDA-6614-tab/tab/ukhls/j_indresp.tab",
            col_select = c(pidp, j_sex_dv, j_age_dv, j_sclfsato))
w11 <- vroom("UKDA-6614-tab/tab/ukhls/k_indresp.tab",
            col_select = c(pidp, k_sex_dv, k_age_dv, k_sclfsato))

w9_11 <- w9 %>% 
  full_join(w10, by="pidp") %>%
  full_join(w11, by="pidp")
w9_11

?pivot_longer

w9_11 <- w9_11 %>%
  pivot_longer(cols = i_sex_dv:k_sclfsato, names_to ="variable", values_to="value") %>%
  separate(variable, into=c("wave", "variable"),sep="_",extra="merge") %>% #into = new col names,
  #variable = names_to #sep=seperator #extra=what to do with extra bits
  pivot_wider(names_from=variable, values_from = value)

#Now we have a person/wave
```

2. Add the variable for lag life satisfaction (life satisfaction in previous wave).

```{r q2}
w9_11 <- w9_11 %>%
  group_by(pidp) %>%
  mutate(laglife = lag(sclfsato)) 
# window functions allow you to work something out without having to reshape data

# find mean for each pidp (person)
w9_11 %>%
  group_by(pidp) %>%
  summarise(mean= mean(sclfsato, na.rm=T)) 
```


3. Reshape the data back to the wide format. Produce a table with the distribution for the number of waves individuals participated in.


```{r q3}

# go longer before wider
w9_11 %>%
pivot_longer(cols = sex_dv:sclfsato, names_to ="variable", values_to="value") %>%
  unite(wave:variable, col="variable") %>%
  pivot_wider(names_from =variable, values_from=value)

```

4. Join the **indresp** data from all 11 waves of the Understanding Society and reshape to the long format. Only keep the variables for pidp, sex, age, region (gor_dv) and life satisfaction.


```{r q4}
# create a vector with the file names and paths
files <- dir(
  # Select the folder where the files are stored.
  "UKDA-6614-tab/tab/ukhls",
  # Tell R which pattern you want present in the files it will display.
  pattern = "indresp",
  # And finally want R to show us the entire file path, rather than just
  # the names of the individual files.
  full.names = TRUE)

files

vars <- c("sex_dv", "age_dv", "gor_dv", "sclfsato")
vars

for (i in 1:11) {
  # Create a vector of the variables with the correct prefix.
  varsToSelect <- paste(letters[i], vars, sep = "_")
  # Add pidp to this vector (no prefix for pidp)
  varsToSelect <- c("pidp", varsToSelect)
  # Now read the data. 
  data <- vroom(files[i], col_select = varsToSelect)
  if (i == 1) {
    all11 <- data  
  }
  else {
    all11 <- full_join(all11, data, by = "pidp")
  }

} 

all11

all11long <- all11 %>%
  pivot_longer(cols = a_sex_dv:k_sclfsato, names_to ="variable", values_to="value") %>%
  separate(variable, into=c("wave", "variable"),sep="_",extra="merge") %>% 
  pivot_wider(names_from=variable, values_from = value)

all11long

#calculate the number of missed waves for each individual (number of NA characters for sex/age)
all11long %>%
  group_by(pidp) %>%
  summarise(nMissed = sum(is.na(sex_dv))) %>%
  mutate(nParticipated = 11-nMissed) %>%
  count(nParticipated)

```

5. Calculate the average life satisfaction by wave and region. Present as a table with regions in rows and waves in columns.

```{r q5}
all11long %>%
  group_by(wave, gor_dv) %>%
  summarise(
    meanLS = mean(sclfsato, na.rm=T)
  ) %>%
  pivot_wider(names_from=wave,values_from=meanLS)
```