---
title: "Week 2. Tasks/exercises for 25 January 2022"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vroom)
```

## The Understanding Society questions

1. What is the difference between **indresp** and **indall** files in UndSoc? What kind of information can you find in each?

```{r q1}

k_indresp <- vroom("UKDA-6614-tab/tab/ukhls/k_indresp.tab")
k_indall <- vroom("UKDA-6614-tab/tab/ukhls/k_indall.tab")

colnames(k_indresp)
colnames(k_indall)

```

1b. What kind of information can you find in **xwavedat** files?

```{r q1b}

xwavedat <- vroom("UKDA-6614-tab/tab/ukhls/xwavedat.tab")

head(xwavedat)
colnames(xwavedat)

```

2. What is the structure of **egoalt** files?

```{r q2}

k_egoalt <- vroom("UKDA-6614-tab/tab/ukhls/k_egoalt.tab")


```

3. In which files can you find the data on education / highest qualification?

4. What kind of information can you find in the **hhresp** and **hhsamp** files?

```{r q4}

k_hhresp <- vroom("UKDA-6614-tab/tab/ukhls/k_hhresp.tab")
k_hhsamp <- vroom("UKDA-6614-tab/tab/ukhls/k_hhsamp.tab")


```

5. What kind of information can you find in the **newborn** files?

6. What is the pidp variable? Why is it important?

7. Tell us about the individual with pidp 280165 (sex, age, where they live, etc.).

```{r q7}

xwavedat %>%
  filter(pidp == 280165)

```


8. Explore the structure of the questionnaires for wave 11 of UndSoc. 



