---
title: "Week 3. Tasks/exercises for 1 February 2022"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vroom)
```

## Data transformation

1. Open the **indresp** file from week 11. Restrict your sample to women aged 18 to 35.

```{r q1}
k_indresp <- vroom("UKDA-6614-tab/tab/ukhls/k_indresp.tab")

# Look at dist of data - important because of -9 etc.
k_indresp %>% 
  count(k_sex_dv)
k_indresp %>% 
  count(k_age_dv)

#Only women and 18-35 inclusive
filter(k_indresp, k_sex==2, k_dvage>=18 & k_dvage<=35)

#Alternatives
k_indresp %>%
  filter(k_sex_dv ==2 & k_age_dv %in% 18:35)

k_indresp %>%
  filter(k_sex_dv ==2) %>%
  filter(k_age_dv %in% 18:35)

## Not women aged 18-35 (so still women >35)
k_indresp %>%
  filter(!(k_sex_dv ==2 & k_age_dv %in% 18:35))

## Not women or aged 18-35
k_indresp %>%
  filter(!(k_sex_dv ==2 | k_age_dv %in% 18:35))

## Select women aged 18 to 20 and men aged 51 to 53
k_indresp %>%
  filter(k_sex ==2 & k_dvage %in% 18:20| k_sex ==1 & k_dvage %in% 51:53) %>%
  count(k_sex, k_dvage)

```

2. Recode sex in such a way that women are shown as 1 and men are shown as 0.

```{r q2}

## First way - ifelse (give new name to new sex variable)
k_indresp %>%
  mutate(sex2 = ifelse(k_sex == 1, 0,
                       ifelse(k_sex == 2, 1, NA))) %>%
  count(sex2)

## Alternative - case_when - assigns everything else NA - give new name to new sex variable

k_indresp %>%
  mutate(sex1 = case_when(
    k_sex_dv ==1 ~ 0,
    k_sex_dv ==2 ~ 1)
  ) %>%
  count(sex1)
```

3. Recode age into the following age groups:  < 25, 26-45, 46-65, >65 years old.

```{r q3}
summary(k_indresp$k_dvage)
# -9 = missing
# 3 missing values

k_indresp %>%
  mutate(k_age_dv2 =case_when(
    k_age_dv <0 ~ NA_character_,
    k_age_dv %in%  0:25 ~ "<25",
    k_age_dv %in% 26:45 ~ "26-45",
    k_age_dv %in% 46:65 ~ "46-65",
    k_age_dv >65 ~ ">65"
  )) %>% 
  count(k_age_dv2)

## Create a variable with the following values: Women aged 18 to 20, all other women,
## everyone else

## This shows the importance of the first line of code going down!

k_indresp %>%
  mutate(k_group =case_when(
    k_sex_dv ==2 & k_age_dv %in% 18:20 ~ "women 18-20",
    k_sex_dv ==2 ~ "other women",
    TRUE ~ "everyone else"
  )) %>%
  count(k_group)


```

4. Create a variable with the following values: "young women" (women aged 16 to 35), "young men" (men aged 16 to 35),
"older women" (women aged 35 to 60), "older men" (men aged 35 to 60). Other cases should be shown as missing. 

```{r q4}
k_indresp %>%
  mutate(k_age_sex_dv3 =case_when(
    k_sex_dv ==2 & k_age_dv %in% 16:35 ~ "young women",
    k_sex_dv ==1 & k_age_dv %in% 16:35 ~ "young men",
    k_sex_dv ==2 & k_age_dv %in% 35:60 ~ "older women",
    k_sex_dv ==1 & k_age_dv %in% 35:60 ~ "older men"
  )) %>%
  count(k_age_sex_dv3)

```

5. Create a data frame showing the mean and median net income and its standard deviation.

```{r q5}
k_indresp %>%
count(k_fimnnet_dv)

## Pull out column as vector
k_indresp %>%
  pull(k_fimnnet_dv) %>% summary()

k_indresp %>%
  pull(k_fimnnet_dv) %>% hist(breaks=1000)

k_indresp %>%
  summarise(
    mean = mean(k_fimnnet_dv, na.rm =TRUE),
    median = median(k_fimnnet_dv, na.rm=T),
    sd =sd(k_fimnnet_dv, na.rm=T))

```

6. Summarise mean income by sex. Summarise mean income by sex and age group.


```{r q6}

# sex
k_indresp %>%
  mutate(sex1 = case_when(
    k_sex_dv ==1 ~ "men",
    k_sex_dv ==2 ~ "women")
  ) %>%
  mutate(k_age_dv2 =case_when(
    k_age_dv <0 ~ NA_character_,
    k_age_dv %in%  0:25 ~ "<25",
    k_age_dv %in% 26:45 ~ "26-45",
    k_age_dv %in% 46:65 ~ "46-65",
    k_age_dv >65 ~ ">65"
  )) %>% 
  filter(!is.na(sex1)) %>%
group_by(sex1) %>%
  summarise(
    mean = mean(k_fimnnet_dv, na.rm =TRUE),
    median = median(k_fimnnet_dv, na.rm=T),
    sd =sd(k_fimnnet_dv, na.rm=T))

#sex and age
k_indresp %>%
  mutate(sex1 = case_when(
    k_sex_dv ==1 ~ "men",
    k_sex_dv ==2 ~ "women")
  ) %>%
  mutate(k_age_dv2 =case_when(
    k_age_dv <0 ~ NA_character_,
    k_age_dv %in%  0:25 ~ "<25",
    k_age_dv %in% 26:45 ~ "26-45",
    k_age_dv %in% 46:65 ~ "46-65",
    k_age_dv >65 ~ ">65"
  )) %>% 
  filter(!is.na(sex1)& !is.na(k_age_dv2)) %>%
group_by(sex1, k_age_dv2) %>%
  summarise(
    mean = mean(k_fimnnet_dv, na.rm =TRUE),
    median = median(k_fimnnet_dv, na.rm=T),
    sd =sd(k_fimnnet_dv, na.rm=T))
```



