---
title: "Week 3. Tasks/exercises for 1 February 2022"
output:
  github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(vroom)
```

## Data transformation

1. Open the **indresp** file from week 11. Restrict your sample to women aged 18 to 35.

```{r q1}
k_indresp <- vroom("UKDA-6614-tab/tab/ukhls/k_indresp.tab")
```

2. Recode sex in such a way that women are shown as 1 and men are shown as 0.

```{r q2}
```

3. Recode age into the following age groups:  < 25, 26-45, 46-65, >65 years old.

```{r q3}


```

4. Create a variable with the following values: "young women" (women aged 16 to 35), "young men" (men aged 16 to 35),
"mature women" (women aged 35 to 60), "mature men" (men aged 35 to 60). Other cases should be shown as missing. 

```{r q4}


```

5. Create a data frame showing the mean and median net income and its standard deviation.

```{r q5}


```

6. Summarise mean income by sex. Summarise mean income by sex and age group.


```{r q6}


```



